---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(gridExtra)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data, echo=FALSE}
#load("movies.Rdata")
load("~/Documents/Stat_Duke_Univ/w3_Linear_Regression/projectw3/movies.Rdata")
```



* * *

## Part 1: Data

```{r Exploring Data, warning =FALSE}
# Relationship: critics_score (Rotten Tomatoes) vs. imdb_rating (IMDB)
summary(movies$critics_score)
summary(movies$audience_score)
summary(movies$imdb_rating)
summary(movies$imdb_num_votes)
summary(movies$imdb_num_votes)
```

#### Histograms

Only the IMDB rating looks like a normal ditribution.
On the other hand Rotten Tomatoes critics and user gives more often high scores than lower. The distributions does not look normal at all 

```{r Plot1, echo=FALSE}
require(gridExtra)
# Histogram critics_score 
p1<-ggplot(movies,aes(critics_score))+
  geom_histogram(binwidth = 10,fill="lightblue", colour="black")+
  theme(axis.text=element_text(size=8),
  axis.title=element_text(size=8,face="bold")) +    # angle = 0
  theme(legend.text = element_text(size = 8, colour = "darkgrey", angle = 0)) +
  theme(axis.text = element_text(colour = "black", angle=0)) + 
  xlim(0,100)+
  scale_fill_brewer(palette="Paired" )
# Histogram audience_score
p2<-ggplot(movies,aes(audience_score))+
  geom_histogram(binwidth = 10,fill="lightblue", colour="black")+
  theme(axis.text=element_text(size=8),
  axis.title=element_text(size=8,face="bold")) +    # angle = 0
  theme(legend.text = element_text(size = 8, colour = "darkgrey", angle = 0)) +
  theme(axis.text = element_text(colour = "black", angle=0)) + 
  xlim(0,100)+
  scale_fill_brewer(palette="Paired" )
# Histogram imdb_rating
p3<-ggplot(movies,aes(imdb_rating))+
  geom_histogram(binwidth = 0.5,fill="lightblue", colour="black")+
  theme(axis.text=element_text(size=8),
  axis.title=element_text(size=8,face="bold")) +    # angle = 0
  theme(legend.text = element_text(size = 8, colour = "darkgrey", angle = 0)) +
  theme(axis.text = element_text(colour = "black", angle=0)) +  
  xlim(0,10)+
  scale_fill_brewer(palette="Paired" )

# Histogram imdb_num_votes
movies<-movies%>%mutate(LnNumVotes=log(imdb_num_votes))
p4<-ggplot(movies,aes(LnNumVotes))+
  geom_histogram(binwidth = 1.5,fill="lightblue", colour="black")+
  theme(axis.text=element_text(size=8),
  axis.title=element_text(size=8,face="bold")) +    # angle = 0
  theme(legend.text = element_text(size = 8, colour = "darkgrey", angle = 0)) +
  theme(axis.text = element_text(colour = "black", angle=0)) +  scale_fill_brewer(palette="Paired" ) + xlim(0,16)

grid.arrange(p1, p2, p3, p4, ncol=2)
```

```{r IMDB Tomatoes number of votes linear relationship, warning =FALSE}
p2<-ggplot(data = movies, aes(x =imdb_rating, y = audience_score)) +
geom_jitter()+geom_smooth(method = "lm", se = FALSE) + ylim(-5,105) + xlim(-0.5,9.5)
# scatter plot critics_score (Tomatoes) vs. imdb_rating 
p1<-ggplot(data = movies, aes(x =imdb_rating, y = critics_score)) +
geom_jitter()+geom_smooth(method = "lm", se = FALSE) + ylim(-5,105) + xlim(1,9.5)
# scatter plot imdb_num_votes ~ imdb_rating
p3<-ggplot(data = movies, aes(x= log(imdb_num_votes), y= imdb_rating)) +
geom_jitter()+geom_smooth(method = "lm", se = FALSE)  + xlim(4.5,15) + ylim(-0.5,10.5)
#Scatter plot imdb_num_votes~critics_score(Tomatoes)
p4<-ggplot(data = movies, aes(x= log(imdb_num_votes), y= audience_score)) +
geom_jitter()+geom_smooth(method = "lm", se = FALSE)  + xlim(4.5,15) + ylim(-5,105)
grid.arrange(p2, p1, p3,p4, ncol=2)

```

```{r slopes}
m<-lm(audience_score~imdb_rating, data=movies) ; m$coefficients[2]
sm.m<-summary(m); sm.m$coefficients[2,]
m<-lm(critics_score~imdb_rating, data=movies) ; m$coefficients[2]
sm.m<-summary(m); sm.m$coefficients[2,]
m<-lm(imdb_rating~log(imdb_num_votes), data=movies) ; m$coefficients[2]
sm.m<-summary(m); sm.m$coefficients[2,]
```


In the context of the relationship between the critics' score (Rotten Tomatoes) and the IMDB rating the slope tell us that:

**For each additional point on rating scale of the IBMD, the model predicts 16 more points of the Rotten Tomatoes audience score, on average.**

**For each additional point on rating scale of the IBMD, the model predicts 20 more points of the Rotten Tomatoes critics score, on average.**

**verytime that the number of voters is multiplied by 10, the model predicts 0.1 more points of the IMDB rating, on average.**

$$scale _x:0-10$$
$$scale_y=0-100$$
This means that with our model when IMDB rating is 3.6 points the model predicts 0 points for Rotten Tomatoes and for 8.6 IMDB rating points the model gives the maximum of possible critics score (100). The IMDB rating's interval is narrower.

The Rotten Tomatoes critics seems more generous than the IMDB critics

```{r linear}
m1<-lm(formula = critics_score ~ imdb_rating,data=movies)
Sm.m1=summary(m1)
Sm.m1      # Sm.m1$coefficients

m2<-lm(formula = audience_score ~ imdb_rating,data=movies)
Sm.m2=summary(m2)
Sm.m2 

m3<- lm(formula = imdb_num_votes ~ imdb_rating,data=movies)
Sm.m3=summary(m3)
Sm.m3   
```


Films with more number of votes (IMDB)
```{r Most voted}
x<-movies[movies$imdb_num_votes >= 600000,]
z=names(movies)
x<-movies[movies$imdb_rating > 4 & movies$top200_box == "yes",]
x<-movies[movies$top200_box == "yes",]
x<-x[order(-x$imdb_rating),]

my_var=names(movies)
my_var<-my_var[c(1,13,14,16,18,15,17,19,20,23,24)]
# Best movie Winner & Nomination
x<-movies[movies$best_pic_win == "yes" & movies$best_pic_nom == "yes",]
x<-x[order(-x$imdb_rating),]
xx=x[my_var]
xx[,c(1:2,4:5,11)]
summary(xx)
# Top 200 movies Top200_Box
x<-movies[movies$top200_box == "yes",]
x<-x[order(-x$imdb_rating),]
xx=x[my_var]
xx[,c(1:2,4,6,8)]
summary(xx)
# Highest imdb rating & Certified fresh
x<-movies[movies$imdb_rating >= 8 & movies$critics_rating == "Certified Fresh",]
x<-x[order(-x$imdb_rating),]
xx=x[my_var]
xx[,c(1:2,4,6,8,11)]
summary(xx)

x<-movies[movies$best_pic_nom == "yes" & movies$dire == "Certified Fresh",]

# Oscar nomination Best Pictutre
x<-movies %>% filter(best_pic_nom == "yes" )

x<-x[order(-x$imdb_rating),]
xx=x[my_var]
xx[,c(1:2,4,6,8,11)]
 # Director
a<-data.frame(table(subset(x,select=c(title,director,actor1))))
a[a$Freq >=1,]
#Actor frequency
a<-data.frame(table(subset(x,select=c(actor1))))
a[a$Freq >=2,]
a<-data.frame(table(subset(x,select=c(actor2))))
a[a$Freq >=2,]


# Table actors
table(subset(movies,select=c(actor1)))
my_var=c(movies$title,movies$imdb_rating,movies$imdb_rating,movies$critics_score,movies$audience_score,movies$critics_rating,movies$audience_rating)
#x[,my_var]
my_var

a<-data.frame(table(subset(movies,select=c(actor1))))
a[a$Freq > 3,]

a<-data.frame(table(subset(movies,select=c(actor2))))
a[a$Freq > 2,]

a<-data.frame(table(subset(movies,select=c(director,best_pic_nom))))

a[a$Freq >= 2,]
```

* * *

## Part 2: Research question

### 1.  What attributes make a movie popular 
### 2.  She is also interested in learning something new about movies


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 4: Modeling

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

```{r Normal distribution}
mu=mean(movies$imdb_rating); sig=sd(movies$imdb_rating)
nIMDB=length(movies$imdb_rating)
x   <- seq(1,10,length=nIMDB)
normalIMDB  <- dnorm(x,mean=mu, sd=sig)
plot(x,normalIMDB, type="l", lwd=1, xlab="IMDB Rating", ylab="",
  main="Normal Distribution")
d <- density(movies$imdb_rating) # returns the density data 
plot(d) # plots the results
p1<-ggplot(movies, aes(imdb_rating)) +
  geom_density(colour="blue")


normal<-rnorm(651,mean=mu, sd=sig)

# Plot 3 Desities Distribution: IMDB,NORMAL (centered at IMDB mean,same sd), Rotten Tomatoes (Rescaled to 10)  
set.seed(1234)
Who<-c(rep("IMDB",651),rep("RottenT-Scale10",651), rep("Normal",651))
normalIMDB=round(rnorm(651,mean=mu, sd=sig),1)
rating<-data.frame(c(movies$imdb_rating,movies$critics_score/10, normalIMDB))
critics=data.frame(cbind(Who,rating))
names(critics)=c("who","rating")
rating.mean=c(mean(movies$imdb_rating),mean(movies$critics_score/10), mean(normalIMDB))
ggplot(critics, aes(x=rating, fill=who)) + geom_density(alpha=.2) +
    geom_vline(data=critics, aes(xintercept=rating.mean[1]), colour="red",
               linetype="dashed", size=1) + 
    geom_vline(data=critics, aes(xintercept=rating.mean[2]), colour="blue",
               linetype="dashed", size=1) +
    geom_vline(data=critics, aes(xintercept=rating.mean[3]),  colour="red",
               linetype="dashed", size=1)
movies<-data.frame(cbind(movies, normal.IMDB=normalIMDB))
p1<-ggplot(movies, aes(x=genre, y=imdb_rating, fill=genre)) + geom_boxplot()+ ylim(1,10)
p2<-ggplot(movies, aes(x=genre, y=critics_score/10, fill=genre)) + geom_boxplot()+ ylim(1,10)
grid.arrange(p1, p2, ncol=2)

p1<-ggplot(data = movies, aes(x =normal.IMDB, y = imdb_rating, colour=genre)) +
geom_jitter()+geom_smooth(method = "lm", se = FALSE) + ylim(1,10) + xlim(1,10)
p1
rm(p1)
# There is NO CORRELATIOIN between the IMDB and a normal population centered at IMDB_rating mean and with a standard variation equal to the standard variation of IMDB_rating
m1<-lm(formula = imdb_rating ~ normalIMDB,data=movies)
Sm.m1=summary(m1)
Sm.m1
```

